---
title: "Misc notes"
editor: source
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE}
source("./R/init.R")
log_info("Called misc-notes notebook")
```

## Thought 1

What is the effect of treating a patient with joint infection with a revision surgical procedure compared to a less invasive procedure, known as dair?
What about the effect of treating with long relative to short duration antibiotic where the definition (and action) of long vs short duration are different within each level (i.e. a nested design rather than crossed).
The usual experimental approach here is to apply random assignment of all possible treatment combinations, which with two levels 



Imagine a situation where we have a nested factorial design with factors A (control vs active) and B (control vs active).
Contrary to the usual situation, we assume that active for A comes in two flavours, active-1 and active-2.
Usually, we assume that the control vs active for B are different under each level of A.
Here, we assume that control vs active for B are different under each of the possibilities for A.




trial with control vs treatment with a single dichotomous covariate in the model (X) and where the outcome is death by day 30. 
We specify a conditional logistic regression model with linear predictor $\eta = \beta_0 + \beta_1 I(trt_i == 1) + \beta_2 I(sex = male)$. 
By chance, all the control group are male and all the treatment group are female. 
Logically, we would not be able to separate the effect of treatment from sex, but model it anyway.

```{r, eval = F}
N <- 10000
d <- data.table(id = 1:N)
d[, trt := rbinom(.N, 1, 0.5)]
d[trt == 0, sex := 1]
d[trt == 1, sex := 0]
beta_0 <- qlogis(0.7)
beta_trt <- log(2)
beta_sex <- log(0.2)
d[, eta := beta_0 + beta_trt * trt + beta_sex * sex]
d[, y := rbinom(.N, 1, prob = plogis(eta))]
head(d)
```

Basic logistic regression.

```{r, eval = F}
#| class-output: stan
#| echo: false
cat(readLines("stan/logistic-demo-01.stan"), sep = "\n")
```

Fit the model and extract posterior.

```{r, eval = F}
m1 <- cmdstanr::cmdstan_model("stan/logistic-demo-01.stan")

d_grp <- d[, .(y = sum(y), n = length(y)), keyby = .(trt, sex)]
ld <- list(
  N = nrow(d_grp), 
  y = d_grp$y, n = d_grp$n, trt = d_grp$trt, sex = d_grp$sex
)
f1 <- m1$sample(
  ld, iter_warmup = 1000, iter_sampling = 2000,
  parallel_chains = 2, chains = 2, refresh = 0, show_exceptions = F, 
  max_treedepth = 13)

post <- data.table(f1$draws(variables = c(
  "b0", "b_trt", "b_sex"
), format = "matrix"))

```

Summarise the posterior draws:

```{r, eval = F}
d_tbl <- melt(post[, 2:3], measure.vars = names(post)[2:3])
d_tbl <- d_tbl[, .(mu = mean(value), 
          sd = sd(value),
          q_025 = quantile(value, prob = 0.025), 
          q_975 = quantile(value, prob = 0.975)), by = variable]

b_tru <- c(beta_trt, beta_sex)
names(b_tru) <- c("b_trt", "b_sex")
d_tbl[, b_tru := b_tru[variable]]
d_tbl
```

So, the posterior mean is nowhere near our known true value that we simulated the data with.

<!-- The model addresses the question of what is the value of knowing sex after knowing treatment assignment. -->
<!-- Given that all those treated were female and all those under control, there is no value in the additional covariate. -->

<!-- The joint posterior of the two effects parameters makes this more clear. -->

```{r, eval = F, echo = F}
#| label: fig-misc1
#| fig-cap: 'Joint posterior for treatment and sex effects.'
#| fig-width: 3.5
#| fig-height: 3.5
#| fig-dpi: 300

d_fig <- copy(post)

# from pt to mm : mm = pt / ggplot2::.pt

ggplot2::theme_update(text = element_text(size = 18 / ggplot2::.pt))
ggplot2::theme_update(axis.text.x = element_text(size = 18 / ggplot2::.pt))
ggplot2::theme_update(axis.text.y = element_text(size = 18 / ggplot2::.pt))
ggplot2::theme_update(legend.position = "top")

ggplot(d_fig, aes(x = b_trt, y = b_sex)) +
  geom_point(alpha = 0.2)
```

<!-- The linear predictor is actually this: -->

<!-- \begin{aligned} -->
<!-- \eta_i = \beta_0 + \beta_{trt} trt_i + \beta_{sex} (1 - trt_i) -->
<!-- \end{aligned} -->

<!-- So, for those that receive control, the contribution to the likelihood is $\eta_i = \beta_0 + \beta_{sex}$. -->
<!-- And for those that receive treatment, the contribution to the likelihood is $\eta_i = \beta_0 + \beta_{trt}$. -->
<!-- In aggregate, we can recover our true log odds of response in  -->

```{r, eval = F, echo = F}
d_fig[, eta_0 := b0 + b_sex]
d_fig[, eta_1 := b0 + b_trt]

d_tbl <- melt(d_fig[, 4:5], measure.vars = names(d_fig)[4:5])
d_tbl <- d_tbl[, .(mu = mean(value), 
          sd = sd(value),
          q_025 = quantile(value, prob = 0.025), 
          q_975 = quantile(value, prob = 0.975)), by = variable]

eta_tru <- c(beta_0 + beta_sex, beta_0 + beta_trt)
names(eta_tru) <- c("eta_0", "eta_1")
d_tbl[, eta_tru := eta_tru[variable]]
d_tbl

```


<!-- Now consider the setting where we compare dair vs review with a covariate for  -->