---
title: "Reminders"
date: today
date-modified: last-modified
editor: source
editor_options: 
  chunk_output_type: console
sidebar: false
navbar: true
# this is necessary to remove an unwanted boarder around a tabset
include-in-header:
  - text: |
      <style>
      .panel-tabset > .nav-tabs,
      .panel-tabset > .tab-content {
        border: none;
      }
      </style>
---

```{r, echo = FALSE}
source("./R/init.R")
log_info("Called reminders notebook")
```

Just some brief notes/examples to act as basics stats reminders.

## Example - LOTP/LOTE

Assume the following true joint distribution for some outcome $Y$ and its distribution conditional on sex, smoking and vegetarianism.

```{r}
set.seed(1)
d_tru <- CJ(sex = 0:1, smk = 0:1, veg = 0:1)
g <- function(sex, smk, veg){
  -1 + 2 * sex - 1 * smk + 3 * veg +
    -1 * sex * smk - 3 * sex * veg + 1 * smk * veg +
    0.5 * sex * smk * veg
}
# probability of group membership
q <- rnorm(nrow(d_tru))
d_tru[, p_grp := exp(q)/sum(exp(q))]
# probability of outcome
d_tru[, p_y := plogis(g(sex,smk,veg))]
d_tru
```

Take a random sample from the population:

```{r}
n <- 1e7
i <- sample(1:nrow(d_tru), n, T, prob = d_tru$p_grp)
d <- d_tru[i]
d[, y := rbinom(.N, 1, p_y)]
```

Recover the distribution of the covariates (the `p_grp`)

```{r}
# recover distribution of covariates
d[, .(.N/nrow(d)), keyby = .(sex, smk, veg)]
```

Recover the joint distribution of the outcome (the `p_y`)

```{r}
# recover distribution of outcome
d[, .(mu_y = mean(y)), keyby = .(sex, smk, veg)]
```

Estimate the probability of outcome by sex

```{r}
d[, .(mu_y = mean(y)), keyby = sex]
```

We are interested in $\mathbb{E}[y | sex]$.
Can we get to this via iterated expectation/LOTE?

Think - is all we need LOTP (with extra conditioning), e.g.

$$
\begin{aligned}
Pr(y | sex = s) &= Pr(y | sex = s, smk = 0, veg = 0) Pr(smk = 0, veg = 0|sex = s) + \\
 & \quad Pr(y | sex = s, smk = 0, veg = 1) Pr(smk = 0, veg = 1|sex = s) + \\
 & \quad Pr(y | sex = s, smk = 1, veg = 0) Pr(smk = 1, veg = 0|sex = s) + \\
 & \quad Pr(y | sex = s, smk = 1, veg = 1) Pr(smk = 1, veg = 1|sex = s) 
\end{aligned}
$$ {#eq-mn-1}

Operationalised, based on the known distributions, we get:

```{r}
c(
  "Pr(Y | sex = 0)" = sum(d_tru[sex == 0, p_y] * d_tru[sex == 0, p_grp / sum(p_grp)] ),
  "Pr(Y | sex = 1)" = sum(d_tru[sex == 1, p_y] * d_tru[sex == 1, p_grp / sum(p_grp)] )
)
```

And which can be estimated from the simulated data:

```{r}
c(
  "Pr(Y | sex = 0)" = 
    sum(d[sex == 0, mean(y)] * 
          d[sex == 0, .(p_grp = .N/nrow(d[sex == 0])), keyby = .(smk, veg)]$p_grp ),
  "Pr(Y | sex = 1)" = 
    sum(d[sex == 1, mean(y)] * 
          d[sex == 1, .(p_grp = .N/nrow(d[sex == 1])), keyby = .(smk, veg)]$p_grp)
)
```

Take @eq-mn-1, multiply both sides by $y$ and then sum over all $y$

$$
\begin{aligned}
\sum_y y Pr(y | sex = s) &= \sum_y y Pr(y | sex = s, smk = 0, veg = 0) Pr(smk = 0, veg = 0|sex = s) + \\
& \quad \sum_y y Pr(y | sex = s, smk = 0, veg = 1) Pr(smk = 0, veg = 1|sex = s) + \\
& \quad \sum_y y Pr(y | sex = s, smk = 1, veg = 0) Pr(smk = 1, veg = 0|sex = s) + \\
& \quad \sum_y y Pr(y | sex = s, smk = 1, veg = 1) Pr(smk = 1, veg = 1|sex = s)  
\end{aligned} 
$$ {#eq-mn-2}

which (I think) can be re-stated as

$$
\begin{aligned}
\mathbb{E}(Y|sex=s) &= \mathbb{E}(Y|sex = s, smk = 0, veg = 0) Pr(smk = 0, veg = 0|sex = s) + \\
& \quad \mathbb{E}(Y|sex = s, smk = 0, veg = 1)  Pr(smk = 0, veg = 1|sex = s)  + \\
& \quad \mathbb{E}(Y|sex = s, smk = 1, veg = 0)  Pr(smk = 1, veg = 0|sex = s)  + \\
& \quad \mathbb{E}(Y|sex = s, smk = 1, veg = 1)  Pr(smk = 1, veg = 1|sex = s)  \\
&= \sum_{smk,veg} \mathbb{E}(Y|sex = s, smk, veg) Pr(smk, veg|sex = s)
\end{aligned} 
$$ {#eq-mn-3}

also note^[For the proof, consider $P(Y,X|Z) = \frac{Pr(X,Y,Z)}{Pr(Z)}$, $P(Y|X,Z) = \frac{Pr(X,Y,Z)}{Pr(X,Z)}$ and $Pr(X|Z) = \frac{Pr(X,Z)}{Pr(Z)}$.] (just in case one is easier to determine than another):

$$
Pr(smk, veg|sex) = Pr(smk|veg, sex) Pr(veg | sex)
$$

so (perhaps) alternatively:

$$
\begin{aligned}
\mathbb{E}(Y|sex=s) &= \sum_{smk,veg} \mathbb{E}(Y|sex = s, smk, veg) Pr(smk|veg, sex = s) Pr(veg | sex = s)
\end{aligned} 
$$ {#eq-mn-3}




