---
title: "Model specification"

---

```{r, echo = FALSE}
source("./R/init.R")
log_info("Called model-spec notebook")
```

The primary outcome is treatment success at 12 months post platform entry, defined as **all of**: 

1. Alive
2. Clinical cure (no clinical or microbiological evidence of infection)
3. No ongoing use of antibiotics for the index joint; and
4. The prosthesis present after the initial management strategy is complete (destination prosthesis) still in place

hereafter referred to as 'treatment success'.

For each silo and site of infectioin we model the probability of treatment success as follows:

\begin{aligned}
\log{ \left( \frac{\pi_i}{1 - \pi_i} \right) } = \alpha_{s(i),u(i)} + \zeta_i
\end{aligned}

where 

+ $\alpha_{s(i),u(i)}$ denotes the log-odds of treatment success for patient $i$ in silo $s$ and having infection site $u$ in those randomised to the reference groups in the surgery (A), duration (B) and antibiotic type (C) domains
+ $\zeta_i$ is a series of other terms that are context dependent and expanded upon below

### Early infection

Patients with early stage infection are only randomised into the antibiotic type domain; the surgical intervention is fixed to be dair and the antibiotic duration is fixed at 12 weeks.
In other words, there are no treatment comparisons in the surgical nor duration domain for the early stage infection patients; neither do they contribute to the estimation of parameters in the late or chronic silos.

$\zeta$ comprises terms that characterise the effects of non-membership in domains $A$, $B$ and $C$ (60% of the cohort are assumed to enter into domain C irrespetive of their membership in domains $A$ and $B$).
Non-membership parameters are shared across all silos.
$\zeta$ also includes a term to model the effect of treatment allocation within domain $C$.

The $\gamma$ terms denote the effect of non-membership with subscript indicating domain and incorporated into the linear predictor based on an indicator of domain non-membership $\mathbb{I}_{A^\prime(i)}$ for each patient $i$.
In the simulated data, for early stage infection patients, both $\mathbb{I}_{A^\prime(i)}$ and $\mathbb{I}_{B^\prime(i)}$ would evaluate to true for all patients.
While terms for non-membership in domain $C$ are relevant, there are no randomisation options available for early infections in domain $A$ nor $B$.
Therefore, non-membership terms for $A$ and $B$ are not required in the linear predictor for this silo^[In practice, it may be the case that non-membership for domain $B$ is relevant. For example, if an early stage patient receives a one or two-stage surgical procedure.].

For those patients entering into the antibiotic type domain, there is a parameter that charactersises the effect of treatment(s) $\beta_{C,d_{C,j(i)}}$ that is shared across silos^[As such, silo-specific decisions on antibiotic type are not possible.] and for which $d_{C,j}$ is a treatment index.
By definition, this negates the possibility of characterising effect heterogeneity.
The treatment effect is included (or not) in the linear predictor based on an indicator of domain membership $\mathbb{I}_{C(i)}$ for each patient $i$.

\begin{aligned}
\zeta_i = \gamma_{C} \mathbb{I}_{C^\prime(i)} + \beta_{C,d_{C,j(i)}} \mathbb{I}_{C(i)}
\end{aligned}
 
No other terms are included in the model for the purposes of simulation.
In the full model, there would be additional terms to capture time trends, site variation and prognostic covariates.
For $C$, the treatment effects could be partially pooled across silos.
However, note that a population view of the 'average' OR across the silos, which could be obtained by sampling from a normal distribution with variance set to the group level variance posterior, would likely be attenuated towards the null and have more uncertainty relative to a pooled estimate.

### Late infection

Patients with late infection enter all domains, but with some restrictions.
They are randomised to dair vs revision in the surgical domain. 
Patients allocated to revision will receive a one or two-stage procedure, the plan for which needs to be captured at the time of randomisation.
Patients allocated to dair, do not contribute to the estimation of duration effects.
Patients allocated to revision for whom the plan is to undertake a one-stage procedure are allocated to 6 vs 12 weeks following stage one.
Patients allocated to revision for whom the plan is to undertake a two-stage procedure are allocated to 7 days vs 12 weeks following stage two.

$\zeta$ again comprises terms that characterise the effects of non-membership in domains $A$, $B$ and $C$ (60% of the cohort are assumed to enter into domain C irrespetive of their membership in domains $A$ and $B$) along with terms to model the effects of treatment allocation within each domain.
All patients are assumed to enter the surgical domain and so $\gamma_{A}$ can be omitted.
Patients allocated to dair, would contribute to the estimation of $\gamma_{B}$ and $\gamma_{C}$ is as previously defined.
However, the inclusion of $\gamma_{B}$ creates an identifiability issue for $\alpha$ and $\beta_A$ and therefore is not included for the purposes of the simulations.

\begin{aligned}
\zeta_i &= \gamma_{C} \mathbb{I}_{C^\prime(i)} + \beta_{A,s(i),d_{A,j(i)}}\mathbb{I}_{A(i)} + \\
&\qquad  \beta_{B_{1},s(i),d_{B_1,j(i)}}\mathbb{I}_{one-stage(i)}  +  \beta_{B_{2},s(i),d_{B_2,j(i)}}\mathbb{I}_{two-stage(i)} +  \beta_{C,d_{C,j(i)}}\mathbb{I}_{C(i)}
\end{aligned}

+ $\beta_{A,s,d_{A,j}}$ represents a (silo specific) parameter characterising the effect of revision relative to dair 
+ $\beta_{B_{1},s,d_{B_1,j}}$ represents a (silo specific) parameter characterising the effect of 12 weeks relative to 6 weeks in patients who were planned to have a one-stage procedure.
+ $\beta_{B_{2},s,d_{B_2,j}}$ represents a (silo specific) parameter characterising the effect of 12 weeks relative to 7 days (post second stage) in patients who were planned to have a two-stage procedure.
+ $\beta_{C,d_{C,j}}$ is as previously defined, again noting that the parameter is currently pooled over all silos.

The reference level parameters for the treatment effects are all set to zero.
The terms for the $B$ domain are silo-specific, i.e. unpooled.
These should be considered for partial pooling across the late and chronic silos.
However, the fact that one vs two-stage were not randomised for the late silo might lead to some confounding and therefore complicate the interpretation of these terms even further.

### Chronic infection

Patients with chronic infection enter all domains without restrictions.
They are randomised to one vs two-stage in the surgical domain. 
Patients allocated to one-stage are allocated to 6 vs 12 weeks following stage one.
Patients allocated to two-stage are allocated to 7 days vs 12 weeks following stage two.
Entry into the type domain is as above.

$\zeta$ again comprises terms that characterise the effects of non-membership in domains $A$, $B$ and $C$ (60% of the cohort are assumed to enter into domain C irrespetive of their membership in domains $A$ and $B$) along with terms to model the effects of treatment allocation within each domain.
However, since all patients are assumed to enter the surgical and duration domains, $\gamma_{A}$ and $\gamma_{B}$ can be omitted.

\begin{aligned}
\zeta_i &= \gamma_{C} \mathbb{I}_{C^\prime(i)} + \beta_{A,s(i),d_{A,j(i)}}\mathbb{I}_{A(i)} + \\
&\qquad  \beta_{B_{1},s(i),d_{B_1,j(i)}}\mathbb{I}_{one-stage(i)}  +  \beta_{B_{2},s(i),d_{B_2,j(i)}}\mathbb{I}_{two-stage(i)} +  \beta_{C,d_{C,j(i)}}\mathbb{I}_{C(i)}
\end{aligned}

With definitions for the terms as above.

## Interpretation

Some additional more detailed and narrative is perhaps useful on the interpretation of parameters.

As noted previously, the baseline reference groups are defined as follows:

::: {.column width="50%"}

| Silo |  A | B | C | 
|------|-----|---|---|
| early |  - | - | no-rif | 
| late  | dair | - | no-rif |
| chronic | one | 6wk | no-rif |

: Groups corresponding to baseline log-odds of treatment success {.striped .hover}
:::

As such, we have:

+ $\alpha_{[early, \cdot]} \quad$ - this parameter describes the log-odds of treatment success for the sample members with early infection who are randomised within Domain C (that is, were not part of the 40% Domain C exclusion set) to no-rif and had no randomisation options for domains A and B. The parameter makes no adjustment for what surgery a patient received, nor how long their duration of antibiotic was. 

+ $\alpha_{[late, \cdot]} \quad$ - this parameter describes the log-odds of treatment success for the sample members with late infection who were randomised to dair in Domain A, had no randomisation options for Domain B, which is a deterministic consequence of being randomised to dair in Domain A, and were randomised to no-rif in Domain C.

+ $\alpha_{[chronic, \cdot]} \quad$ - this parameter describes the log-odds of treatment success for the sample members with chronic infection who were randomised to one-stage in Domain A, were randomised to 6wksp1 in Domain B and were randomised to no-rif in Domain C. 


```{r, eval = F, echo = F}
p <- c(0.7, 0.5, 0.6)
pr_grp <- c(0.3, 0.5, 0.2)
nsim <- 10000
m <- matrix(NA, ncol = 2, nrow = nsim)

eta_tru <- sum(qlogis(p) * pr_grp)


for(i in 1:nsim){
  d <- data.table(
    ii = sample(1:3, 134, replace = T, prob = pr_grp)
  )
  d[, y := rbinom(.N, 1, prob = p[ii])]
  
  d_mod <- d[, .(y = sum(y), .N), keyby = ii]
  d_mod[, p_obs := y/N]
  d_mod[, eta := qlogis(p_obs)]
  f1 <- glm(cbind(y, N-y) ~ 1, data = d_mod, family = "binomial" )
  s1 <- summary(f1)
  m[i, ] <- c(d_mod[, sum(eta * N)/sum(d_mod$N)], s1$coefficients[1,1])
}

d_fig <- data.table(
  observed = m[, 1],
  estimated = m[, 2],
  true = eta_tru
)

d_fig <- melt(d_fig, measure.vars = names(d_fig))


ggplot(d_fig, aes(x = variable, y = value)) +
  geom_point() +
  geom_hline(data = d_fig[, .(value = mean(value)), keyby = variable], 
             aes(yintercept = value, col = variable, group = variable))
```



::: callout-note
The following are incomplete:
:::

### Late infection

In the surgical domain, we consider the effect (log-odds-ratio) of revision relative to dair.
Within revision, patients may receive one or two-stage procedures.
The log-odds of treatment success for those receiving a one-stage procedure, are randomised to 6wk duration and receive no-rif is $\alpha_{late,\cdot}$.
Similarly, this is true for those receiving a (non-randomised) two-stage procedure, are randomised to 7day duration and receive no-rif.

The effect of revision relative to dair is thus in the context of some weighted average of the response under the one-stage and two-stage procedures.

The model characterises associations between the log-odds of treatment success and each of the parameters.
The effect of each term is considered while holding all others constant.

:::callout-note
Is overlap and extrapolation an issue here???
:::


### Chronic infection


TODO

## Priors

As with the linear predictor, the priors are currently targeted towards the simulation work but may be appropriate for the final model.

### Intercepts

The silo and infection site specific intercepts are given independent normal priors

\begin{aligned}
\alpha_{s(i),u(i)} \sim \mathcal{N}(0, 1.5^2)
\end{aligned}

On the probability scale, these concentrate on 0.5 with 95% of the density between 0.04 and 0.96, approximately uniform across this interval.

While the intercept priors are currently independent, consideration should be given to partial pooling across site of infection.
However, the concept of exchangeability is perhaps dubious for this set of parameters due to the fact that the baseline log-odds of success must be interpreted based on the domain level options available to each group.

For example, there are no randomisation options for early stage infection in domains A and B and therefore the baseline log-odds (intercept) refers to the cohort of patients that were not randomised within A nor B and randomised to the reference group for domain C.
In contrast, for the late stage patients, the baseline log-odds of treatment success is interpreted contextually for the cohort of patients that were randomised to the reference level of domains A, B and C, see below.

::: {.column width="50%"}

| Silo |  A | B | C | 
|------|-----|---|---|
| early |  - | - | no-rif | 
| late  | dair | - | no-rif |
| chronic | one | 6wk | no-rif |

: Groups corresponding to baseline log-odds of treatment success {.striped .hover}
:::

### Domain non-membership

The domain non-membership parameters are given standard normal independent priors.
These parameters are pooled across all silos.

\begin{aligned}
\gamma_\mathcal{D} \sim \mathcal{N}(0, 1)
\end{aligned}

### Interventions

A hierarchical prior structure may be appropriate for some of the intervention effects dependent on what can be assumed regarding exchangeability.

#### Surgical domain

Silo-specific, independent normal priors are assumed for the surgical domain.
There is no pooling, we simply compare dair vs revision in the late stage patients and one vs two stage in the chronic stage patients.

\begin{aligned}
\beta_{A,s,d_{A,j}} \sim \mathcal{N}(0, 1)
\end{aligned}

The reference level is set to 0.

#### Antibiotic duration domain

While a hierarchical structure could be considered for this domain, independent normal priors are currenlty assumed for the duration domain.
Currently, there is no pooling.
We compare 6 vs 12 weeks for late stage patients having a one stage procedure and 7 days vs 12 wks for those having a two stage procedure.
Independently, we compare 6 vs 12 weeks for chronic stage patients having a one stage procedure and 7 days vs 12 wks for those having a two stage procedure.

\begin{aligned}
\beta_{B_k,s,d_{B,j}} \sim \mathcal{N}(0, 1)
\end{aligned}

The reference level is set to 0.

Theoretically, partial pooling could be implemented across the late and chronic silos.

#### Antibiotic type domain

A hierarchical structure could also be considered for this domain.
The 'average' across silos could be obtained by sampling from the posterior assuming normality based on the parameter for the group variance.
With three groups the extent of the partial pooling is going to be heavily influenced by the prior rather than the data.

Currently we adopt a standard normal prior for the treatment effects in this domain, which are completely pooled across silos.

\begin{aligned}
\beta_{C,d_{C,j}} \sim \mathcal{N}(0, 1)
\end{aligned}

The reference level is set to 0.



