---
title: "Model implementation"

---

```{r, echo = FALSE}
source("./R/init.R")
log_info("Called model-implementation notebook")
```



The simplest implementation I could think of split the data into silo-specific chunks, mimicking the formulation in the model specification earlier.
Nevertheless, the model is cumbersome.

```{r}
#| class-output: stan
#| echo: false
cat(readLines("stan/model-sim-02.stan"), sep = "\n")
```

Below I create a data set with many more records that we will have available in order to see if we can recover the parameters.
I fit the above model and then extract and summarise the posterior.

```{r}
m2 <- cmdstanr::cmdstan_model("stan/model-sim-02.stan")

set.seed(1)
ll <- get_trial_data(N = 5e4)
lsd <- get_stan_data(ll$d_i)
ld <- lsd$ld
d_b <- copy(lsd$d_b)

f2 <- m2$sample(
  ld, iter_warmup = 1000, iter_sampling = 2000,
  parallel_chains = 2, chains = 2, refresh = 0, show_exceptions = F, 
  max_treedepth = 13)

post <- data.table(f2$draws(variables = c(
  "eta"
), format = "matrix"))

# create index field
post <- melt(post, measure.vars = names(post))
post[, idx := gsub(".*\\[", "", variable)]
post[, idx := gsub("\\]", "", idx)]
post[, idx := as.integer(idx)]

d_fig <- cbind(d_b, post[, .(eta_med = median(value), 
         eta_q025 = quantile(value, prob = 0.025),
         eta_q975 = quantile(value, prob = 0.975)), keyby = idx])
```

@fig-post-vs-tru shows a comparison between the true log-odds of treatment success with the 95% credible interval obtained from the model.
It suggests a strong association between the true and estimated log-odds of treatment success for this particular dataset.

```{r}
#| label: fig-post-vs-tru
#| fig-cap: 'Scatter plot comparing true vs estimated 95% credible interval for log-odds treatment success.'
#| fig-cap-location: margin
#| fig-height: 6

ggplot(d_fig, aes(x = eta, y = eta_med)) +
  # geom_point() +
  geom_linerange(aes(ymin = eta_q025, ymax = eta_q975)) +
  geom_abline(intercept = 0, slope = 1, lwd = 0.1) +
  scale_x_continuous("True log-odds success") +
  scale_y_continuous("Posterior log-odds success (95% credible interval)") +
  facet_grid(silo ~ joint) +
  ggtitle("True log-odds success vs 95% credible interval")
```

