---
title: "Population structure"
date: today
date-modified: last-modified
---

```{r, echo = FALSE}
source("./R/init.R")
log_info("Called population-structure notebook")
```

The following specification is for the partially factorial structure of the study.
The outcome model is specified separately.


## Silo

The total sample size is divided across the silos in proportion to the values described in the table below.
These proportions are expectations. 
Each simulated dataset will vary somewhat from these proportions due to the stochastic nature of the generation process.

::: {.column width="50%"}

| Silo |  $\pi$ |
|------|-----|
| early |  0.3 |
| late  |   0.5 |
| chronic | 0.2 |

: Silo categories ($\pi$ denotes probability of membership) {.striped .hover}
:::


## Site of infection (joint)

Each silo comprises patients assumed to have primary infection in either knee or hip (not both).
The proportion of infections associated with each joint for each silo are shown below.
As previously, these proportions are expectations and the empirical proportions observed in simulated datasets will vary from these.

::: {.column width="70%"}

Silo | Joint |  $\pi$ |
-----|-----|----|----|
early | knee |  0.4 | 
early |hip |  0.6 | 
late |knee |  0.7 | 
late |hip |  0.3 | 
chronic |knee | 0.5 | 
chronic |hip |  0.5 | 

: Site of infection ($\pi$ denotes probability of site infection conditional on silo membership) {.striped .hover}
:::

## Randomisation into surgery domain (A)

Only the late-stage infection cohort presently have randomisation options for surgery type.
Other silos are assigned by clinician/patient preference.

## Surgery domain (A)

Allocation probabilities (conditional on entry into domain A) are shown below.
Early stage patients do not receive randomisation and are assumed to receive DAIR.
Chronic stage patients do not receive randomisation and are assumed to receive DAIR, one-stage and two-stage based on clinician assessment.

::: {.column width="70%"}

Silo | Surgery type ($a$) |  $\pi$ |
-----|-----|----|----|
early | dair |     - | 
late | dair |     0.5 | 
late |revision |  0.5 | 
chronic |dair |  - | 
chronic | one-stage | - | 
chronic |two-stage |  - | 

: Allocation within domain A ($\pi$ denotes probability allocation to surgery type conditional on silo membership) {.striped .hover}
:::

## Intended surgery

Intended or planned surgery relates to the surgical preference, which may be conditional on which intervention is randomised.

For early stage infection, we take preference towards dair, one and two-stage surgery to be distributed as 90%, 10% and 0%.

For late stage infection, we take preference towards one and two-stage as 30% and 70%, if randomised to revision.
When randomised to dair, we assume the preference to dair, one and two-stage is 20%, 24% and 56%.

For chronic stage infection, we take preference towards dair, one and two-stage surgery to be distributed as 20%, 20% and 60%.

## Randomisation into duration domain (B)

Entry into domain B is dependent on the surgery that was actually received.

Surgery received is dependent on whether surgery was randomised (as in late infection patients) and what the surgical preference is (as in chronic infection patients).

While allocated surgery (dair, one and two-stage) is a linear function of randomised surgery and preference, the surgery received may deviate from the original plan.


## Duration domain (B)

Within the duration domain, allocation is to long vs short duration, the meaning of which differs under one and two-stage surgery.
Patients receiving DAIR are assumed to have 12 wk duration (not randomised).
Default randomisation is 1:1 for long vs short within surgery type.


## Randomisation into antibiotic choice domain (C)

The data generating process assume that 60% of the cohort enter into this domain at random, unrelated to risk factors.
The remainder are held out so that we do not over-estimate the operating characteristics, such as power.
 

## Antibiotic choice domain (C)

The treatment options and allocation probabilities for domain C are 1:1.


## Encoded specification

The above specification is bundled into an R package (`roadmap.data`) for consistent data generation for ROADMAP.

```{r}
#| code-fold: false
roadmap.data::get_pop_spec()
```

The following function simulates the design matrix.

```{r}
#| code-fold: false

roadmap.data::get_design
```

The data generation assumptions imply unique patients groups on which we would observe the outcome.
The outcome is known to be heterogenous across these groups and yet the stated goal is to aggregate measures of effect (odds ratios) across all these groups, e.g. no rif vs rif.
However, the effect of interest is assumed to be obtained from the model parameter that characterises the effect of antibiotic type conditional on the other covariates in the model.

```{r}
#| code-fold: false

d <- roadmap.data::get_design()
head(d)
```

