---
title: "Simulation results 2"
subtitle: "Sequential design with early stopping"

---

```{r, echo = FALSE}
source("./R/init.R")
source("./R/util.R")
log_info("Called simulation-results notebook")
```

The trial is performed sequentially.
Entry into silo-specific (plus surgery specific) parameters is coordinated by triggers for superiority, non-inferiority and futility, defined earlier.

## Equal effects in all treatment arms

```{r}
or_lab <- c("OR 1/2" , "OR 1/1.5", "OR 1", "OR 1.5", "OR 2")
or_lev <- c("1/2", "1/1.5", "1", "1.5", "2")
or_num <- c(1/2, 1/1.5, 1, 1.5, 2)
names(or_lab) <- paste0("v0",c(1, 2, 3, 4, 5))
names(or_num) <- paste0("v0",c(1, 2, 3, 4, 5))

flist <- list.files("data", pattern = "sim02-sc01")

toks <- list()
l <- list()
for(i in 1:length(flist)){
  l[[i]] <- qs::qread(file.path("data", flist[i]))
  toks[[i]] <-  unlist(tstrsplit(flist[i], "[-.]"))
}

d_cfg <- rbindlist(lapply(seq_along(l), function(i){
  tok <- toks[[i]]
  m <- data.table(do.call(cbind, l[[i]]$cfg))
  data.table(cbind(sc = tok[2], v = tok[3], analys = 1:nrow(m), m))
}))

# conversion to numeric
d_cfg[, `:=`(
  nsim = as.numeric(nsim),
  N_pt = as.numeric(N_pt),
  b_r1 = as.numeric(b_r1),
  b_r2 = as.numeric(b_r2),
  w_srp2 = as.numeric(w_srp2),
  b_r1d = as.numeric(b_r1d),
  b_r2d = as.numeric(b_r2d),
  b_f = as.numeric(b_f),
  d_sup = as.numeric(thresh_sup),
  d_ni = as.numeric(thresh_non_inf),
  d_fut = as.numeric(thresh_fut)
  )]

# derive the true effect for surgery based on weight combination
d_cfg[, b_r := b_r1 + w_srp2 * b_r2]
d_cfg[, `:=`(b_r1 = NULL, b_r2 = NULL, w_srp2 = NULL)]

d_tru <- melt(d_cfg[
  , .SD, .SDcols = c("sc", "v", "analys", 
                     "b_r", "b_r1d", "b_r2d", "b_f")], 
  id.vars = c("sc", "v", "analys"), value.name = "lor_tru")


i <- 1
d_dec <- rbindlist(lapply(seq_along(l), function(i){
  tok <- toks[[i]]
  # l[[i]]$d_decision

  d_decision <- copy(l[[i]]$d_decision)
  m <- melt(d_decision, id.vars = c("sim", "analys", "quant"), variable.name = "parname")
  
  # Should be right, but just in case...
  if(any(is.na(m$value))){
    message("Some of the decision values are NA in index ", i, " file ", flist[i])
    m[is.na(value), value := FALSE]
  }
  
  # compute the cumulative instances of a decision being made by sim, each 
  # decision type and by parameter
  m[, value := as.logical(cumsum(value)>0), keyby = .(sim, quant, parname)]
  # summarise by analysis, decision type and parameter
  m <- m[, .(pr_val = mean(value)), keyby = .(analys, quant, parname)]
  # put into wide format
  m <- dcast(m, parname + analys ~ quant, value.var = "pr_val")

  cbind(sc = tok[2], v = tok[3], m)
}))

d_stopping <- rbindlist(lapply(seq_along(l), function(i){
  tok <- toks[[i]]
  # l[[i]]$d_decision

  d_decision <- copy(l[[i]]$d_decision[quant %in% c("sup", "inf", "fut")])
  d_decision[, stopped_all := as.numeric(b_r + b_r1d + b_r2d + b_f == 4)]
  
  # d_early_stop <- d_decision[stopped_all == 1, .SD[1], keyby = .(sim, quant)]
  # d_no_stop <- d_decision[!(sim %in% unique(d_stopped_trials$sim)), .SD[.N], keyby = .(sim, quant)]
  # 
  # d_decision <- rbind(d_early_stop, d_no_stop)

  cbind(sc = tok[2], v = tok[3], d_decision)
}))

d_post_smry_2 <- rbindlist(lapply(seq_along(l), function(i){
  tok <- toks[[i]]
  m <- l[[i]]$d_post_smry_2
  cbind(sc = tok[2], v = tok[3], m)
}))

# grouped participant data 
d_all <- rbindlist(lapply(seq_along(l), function(i){
  tok <- toks[[i]]
  m <- l[[i]]$d_grp
  cbind(sc = tok[2], v = tok[3], m)
}))


```

@tbl-sim-design2-par summarises the setup for the simulated effect sizes (from $\log(1/1.25)$ to $\log(2)$). 
All parameters are simulated to have the same effect size such that all parameters are effective, show no effect or are harmful.
Decision thresholds remain constant throughout the duration of the study.
The margins for non-inferiority and superiority are `r log(1/1.1)` and `r log(1.1)` on the log-odds scale applied across all treatment effect parameters.

```{r}
#| label: tbl-sim-design2-par
#| tbl-cap: 'Parameters used to simulate treatment effects and decision thresholds'
#| column: page
#| out-width: 100%


d_tbl <- d_cfg[, .(v, N_pt, b_r, b_r1d, b_r2d, b_f, thresh_sup, thresh_non_inf, thresh_fut)]

d_tbl[, b_r2d := b_r2d + b_r1d]

d_tbl[, `:=`(
  b_r = exp(b_r),
  b_r1d = exp(b_r1d),
  b_r2d = exp(b_r2d),
  b_f = exp(b_f)
)]

g_tbl <- d_tbl |> gt() |> 
  cols_align(
    columns = everything(),
    align = "center"
  )  |> 
  fmt_number(
    columns = c(b_r, b_r1d, b_r2d, b_f),
    decimals = 3
  ) |>
  tab_spanner(
    label = html("Surgical (D<sub>a</sub>)"),
    columns = c(b_r)
  ) |>
  tab_spanner(
    label = html("Duration (D<sub>b</sub>)"),
    columns = c(b_r1d, b_r2d)
  ) |>
  tab_spanner(
    label = html("Type (D<sub>c</sub>)"),
    columns = c(b_f)
  ) |>
  tab_spanner(
    label = html("Decision thresholds"),
    columns = c(thresh_sup, thresh_non_inf, thresh_fut)
  ) |>
  cols_label(
    v = html("Configuration"),
    b_r = html("revision"),
    b_r1d = html("short<br>(one-stage)"),
    b_r2d = html("short<br>(two-stage)"),
    b_f = html("rif"),
    thresh_sup = html("delta<sub>sup</sub>"),
    thresh_non_inf = html("delta<sub>ni</sub>"),
    thresh_fut = html("delta<sub>fut</sub>")
  ) |>
  tab_style(
    style = list(
      cell_borders(
        sides = c("bottom"), color = "black", weight = px(1), style = "solid"
      )),
    locations = list(
      cells_body(
        columns = everything(),
        rows = N_pt == 2500
      )
    )
  ) |>
  tab_options(
    table.font.size = "80%"
  ) 

g_tbl
```

@fig-pr-design2-1 shows the change in the cumulative probability of declaring each of the decision types for each parameter with changing effects size (odds ratios).

All domains are set so that the treatment effects are all equal, e.g. all set to $\log(2)$ etc.

Note:

1. Superiority decisions^[These have been kept with reference to a clinically important difference of zero as the performance is clearly going to be unacceptable with any non-zero measure.] consider the question of whether the treatment arm is superior to the reference as parameterised in the current model. For example *Is revision superior to dair?"* or *Is short duration superior to long when one-stage revision is received?* etc. Also note that for superiority decisions we currently specify the same decision threshold for all domains and we might want to consider the requirement of differing levels of evidence to support our conclusions, accepting that these would also change the likelihood of making a wrong decision, in a repeat sampling setting.
2. Inferiority (as presently implemented) is simply the mirror of superiority and hence answers the question of whether the reference arm is superior to the treatment arm. However, given that futility is implemented as a mechanism to stop enrolment, the cumulative probability of inferiority is not symmetrical for symmetrical positive and negative effects. Basically, what happens under negative effects is that futility kicks in, which stops enrolment, which means that you do not have as much information to inform an inferiority decision.
3. Non-inferiority decisions are implemented for both directions. That is we address the questions of whether treatment is non-inferior to the reference arm and whether the reference arm is non-inferior to the treatmen arm. In the former and where the treatment arm is simulated as being superior, then non-inferiority is trivially achieved (as noted by Tom and Chris in the last session).
4. Futility considers whether there is negligible chance of establishing superiority. This is not in a predictive sense (i.e. not conditional power) but rather simply a direct assessment of the posterior looking at whether there is less than a 5% chance that the treatment arm is superior. We can be more aggressive and therefore make such decisions earlier (albeit with a greater chance that they would be incorrect) with this by increasing the present decision threshold.


```{r}
#| label: fig-pr-design2-1
#| fig-cap: 'Probability of declaring decision by parameter by effect size (all pars set with same OR).'
#| fig-cap-location: margin
#| fig-height: 8
#| fig-width: 7

# put power by scenario, variant and analysis in long format
d_fig <- melt(d_dec, id.vars = c("sc", "v", "analys", "parname"), variable.name = "quant")
# d_fig <- merge(d_fig, d_pars, by = c("sc", "v", "analys", "parname"))
d_fig[, quant := factor(quant, 
                        labels = c("sup", "inf", "fut", "trt ni ref", "ref ni trt"),
                        levels = c("sup", "inf", "fut", "trt_ni_ref", "ref_ni_trt"))]
# add in the number of pts having reached 12 months post rand by analysis num
d_fig <- merge(d_fig, unique(d_cfg[, .(analys, N_pt)]), by = "analys")


d_fig[, or_tru := or_lab[v]]
d_fig[, or_tru := factor(
  or_tru, labels = or_lab, levels = or_lab)]

fx <- g_fx
names(fx) <- c("rev vs ref", "short vs ref (one)", "short vs ref (two)", "rif vs ref")
d_fig[, parname_lab := factor(
  names(fx[parname]), levels = c("rev vs ref", "short vs ref (one)", "short vs ref (two)", "rif vs ref")
  ) ]



ggplot(d_fig, aes(x = N_pt, y = value, group = quant, col = quant)) +
  geom_point(size = 0.5) +
  geom_line(lwd = 0.4) +
  geom_hline(yintercept = 0.05, lwd = 0.2) +
  ggthemes::scale_colour_tableau(
    "", palette = "Tableau 10",
  type = "regular",
  direction = 1) +
  scale_linetype_discrete("") +
  scale_x_continuous("N (12-months post rand)", guide = guide_axis(angle = 45)) +
  scale_y_continuous("Cumulative probability", breaks = seq(0, 1, by = 0.1)) +
  facet_grid(parname_lab ~ or_tru)
```

@tbl-sim-design2-oc1a provides the same detail as the above figure, but makes it easier to see what the magnitudes of the cumulate probabilities are.

```{r}
#| label: tbl-sim-design2-oc1a
#| tbl-cap: 'Cumulative probability of decision'
#| column: page
#| out-width: 100%

# Widen data so that power is shown by col with each col corresponding to an
# analysis
d_tbl <- d_fig[quant %in% c("sup", "fut", "inf")]
d_tbl <- dcast(d_tbl, parname + or_tru ~ quant + analys, value.var = "value")
names(d_tbl) <- gsub("superiority", "sup", names(d_tbl))
names(d_tbl) <- gsub("inf", "inf", names(d_tbl))
names(d_tbl) <- gsub("futility", "fut", names(d_tbl))
d_tbl <- d_tbl[order(or_tru, parname)]
# d_tbl[, parlab := paste0(parname, " - ", get_effect_label(as.character(parname), do_html = F))]
# d_tbl[, parname := NULL]

g_tbl <- d_tbl |> gt(groupname_col = "parname") |> 
  tab_spanner(
    label = html("Superiority"),
    columns = paste0("sup_", 1:5)
  ) |>
  tab_spanner(
    label = html("Futility"),
    columns = paste0("fut_", 1:5)
  ) |>
  tab_spanner(
    label = html("Inferiority"),
    columns = paste0("inf_", 1:5)
  ) |>
  cols_label(
    or_tru = html("OR (true)"),
    sup_1 = html("500"),
    sup_2 = html("1000"),
    sup_3 = html("1500"),
    sup_4 = html("2000"),
    sup_5 = html("2500"),
    fut_1 = html("500"),
    fut_2 = html("1000"),
    fut_3 = html("1500"),
    fut_4 = html("2000"),
    fut_5 = html("2500"),
    inf_1 = html("500"),
    inf_2 = html("1000"),
    inf_3 = html("1500"),
    inf_4 = html("2000"),
    inf_5 = html("2500")
  ) |>
  tab_style(
    style = cell_borders(
      sides = c("left"),
      weight = px(1)),
    locations = cells_body(
      columns = c(sup_1, fut_1, inf_1)
      )
    ) |>
  tab_style(
    style = list(
      cell_borders(
        sides = c("top", "bottom"), color = "red", weight = px(1), style = "solid"
      )),
    locations = list(
      cells_body(
        columns = everything(),
        rows = or_tru == "OR 1"
      )
    )
  ) |>
  tab_options(
    table.font.size = "80%"
  ) 

g_tbl
```

@fig-est-design2-1 shows the distribution of estimated posterior means by simulated effect size, parameter and sample size.

```{r}
#| label: fig-est-design2-1
#| fig-cap: 'Distribution of posterior means (true log OR shown in red).'
#| fig-cap-location: margin
#| fig-height: 8
#| fig-width: 8

# only interted in the distribution of posterior means for now
d_fig <- d_post_smry_2[!is.na(mu), .(sc, v, sim, par, analys, mu)]
d_fig <- merge(d_fig, unique(d_cfg[, .(analys, N_pt)]), by = "analys")

d_fig[, analys := factor(analys)]
d_fig[, N_pt := factor(N_pt)]

lor_lab <- c("log-OR 1/2" , "log-OR 1/1.5", "log-OR 1", "log-OR 1.5", "log-OR 2")
names(lor_lab) <- paste0("v0", 1:5)
# or_lev <- c("1/2", "1/1.5", "1", "1.5", "2")
lor_num <- log(c(1/2, 1/1.5, 1, 1.5, 2))

d_fig[, lor_lab := lor_lab[v]]
d_fig[, lor_lab := factor(
  lor_lab, 
  levels = lor_lab,
  labels = lor_lab)]

d_fig_2 <- CJ(
  lor_lab = factor(lor_lab, levels = lor_lab, labels = lor_lab),
  par = factor(unique(d_fig$par), levels = unique(d_fig$par))
)
d_fig_2[, lor_num := rep(lor_num, each = length(unique(d_fig$par)))]



p <- ggplot(d_fig, aes(x = N_pt, y = mu)) +
  geom_hline(data = d_fig_2, aes(yintercept = lor_num, group = par), col = 2) +
  geom_boxplot() +
  scale_x_discrete("", guide = guide_axis(angle = 45)) +
  scale_y_continuous("Distribution of posterior mean") +
  facet_grid(par~lor_lab)

suppressWarnings(print(p))
```

TODO - refer to the DSA and evaluate specific questions.


Next figure shows the average number of patients per silo, per domain, per treatment.
How many do not contribute to surgery, how many do not contributed to duration etc.
The next figure shows how many patients contributed to the comparisons.

```{r}
#| label: fig-est-design2-3
#| fig-cap: 'Participants contributing to comparison.'
#| fig-cap-location: margin
#| fig-height: 8
#| fig-width: 8

# TODO

```



## Single effective treatment

```{r, eval = FALSE, echo = F}


flist <- list.files("data", pattern = "sim02-sc02")
toks <- list()
l <- list()
for(i in 1:length(flist)){
  l[[i]] <- qs::qread(file.path("data", flist[i]))
  toks[[i]] <-  unlist(tstrsplit(flist[i], "[-.]"))
}

# cfg - decision thresholds currently static but could be varied over time
d_cfg <- rbindlist(lapply(seq_along(l), function(i){
  tok <- toks[[i]]
  m <- data.table(do.call(cbind, l[[i]]$cfg))
  data.table(cbind(sc = tok[2], v = tok[3], analys = 1:nrow(m), m))
}))

d_pars <- melt(d_cfg[, c("sc", "v", "analys", g_effs), with = F], 
               id.vars = c("sc", "v", "analys"), 
               value.name = "lor_tru", variable.name = "parname")


d_trig <- rbindlist(lapply(seq_along(l), function(i){
  tok <- toks[[i]]
  l[[i]]$d_trig

  m <- melt(l[[i]]$d_trig, id.vars = c("sim", "analys", "quant"), variable.name = "parname")
  # Should be right, but just in case...
  m[is.na(value), value := FALSE]
  m[, value := as.logical(cumsum(value)>0), keyby = .(sim, quant, parname)]
  m <- m[, .(pr_val = mean(value)), keyby = .(analys, quant, parname)]
  m <- dcast(m, parname + analys ~ quant, value.var = "pr_val")

  cbind(sc = tok[2], v = tok[3], m)
}))

d_fig <- melt(d_trig, id.vars = c("sc", "v", "analys", "parname"), variable.name = "quant")
d_fig <- merge(d_fig, d_pars, by = c("sc", "v", "analys", "parname"))
d_fig[, quant := factor(quant, 
                        labels = c("superiority", "futility", "inferiority"),
                        levels = c("sup", "fut", "inf"))]
d_fig <- merge(d_fig, unique(d_cfg[, .(analys, N_pt)]), by = "analys")

d_fig[, or_tru := round(exp(lor_tru), 3)]

d_fig[, or_tru := factor(
  or_tru, 
  labels = c("OR 1/1.25" , "OR 1", "OR 1.25", "OR 1.5", "OR 1.75", "OR 2"),
  levels = c("0.8", "1", "1.25", "1.5", "1.75", "2"))]

setkey(d_fig, sc, v, analys, N_pt)
```

```{r, eval = FALSE, echo = F}
#| label: fig-pr-design2-2
#| fig-cap: 'Probability of declaring decision by parameter by effect size (all pars set with same OR).'
#| fig-cap-location: margin
#| fig-height: 8
#| fig-width: 8

ggplot(d_fig, aes(x = N_pt, y = value, group = quant, col = quant)) +
  geom_point(size = 0.5) +
  geom_line(aes(lty = quant), lwd = 0.3) +
  scale_color_discrete("") +
  scale_linetype_discrete("") +
  scale_x_continuous("Sample size", guide = guide_axis(angle = 45)) +
  scale_y_continuous("Cumulative probability", breaks = seq(0, 1, by = 0.2)) +
  facet_grid(parname~or_tru)
```


@tbl-sim-design2-oc2 provides the same detail as the above figure, but makes it easier to see what the magnitudes of the cumulate probabilities are.

```{r, eval = FALSE, echo = F}
#| label: tbl-sim-design2-oc2
#| tbl-cap: 'Cumulative probability of decision'
#| column: page
#| out-width: 100%

d_tbl <- dcast(d_fig, parname + or_tru ~ quant + analys, value.var = "value")
names(d_tbl) <- gsub("superiority", "sup", names(d_tbl))
names(d_tbl) <- gsub("inferiority", "inf", names(d_tbl))
names(d_tbl) <- gsub("futility", "fut", names(d_tbl))
d_tbl <- d_tbl[order(or_tru, parname)]
d_tbl[, parlab := paste0(parname, " - ", get_effect_label(as.character(parname), do_html = F))]
d_tbl[, parname := NULL]

g_tbl <- d_tbl |> gt(groupname_col = "parlab") |> 
  tab_spanner(
    label = html("Superiority"),
    columns = c("sup_1", "sup_2", "sup_3", "sup_4")
  ) |>
  tab_spanner(
    label = html("Futility"),
    columns = c("fut_1", "fut_2", "fut_3", "fut_4")
  ) |>
  tab_spanner(
    label = html("Inferiority"),
    columns = c("inf_1", "inf_2", "inf_3", "inf_4")
  ) |>
  cols_label(
    or_tru = html("OR (true)"),
    sup_1 = html("1000"),
    sup_2 = html("1500"),
    sup_3 = html("2000"),
    sup_4 = html("2500"),
    fut_1 = html("1000"),
    fut_2 = html("1500"),
    fut_3 = html("2000"),
    fut_4 = html("2500"),
    inf_1 = html("1000"),
    inf_2 = html("1500"),
    inf_3 = html("2000"),
    inf_4 = html("2500")
  ) |>
  tab_style(
    style = cell_borders(
      sides = c("left"),
      weight = px(1)),
    locations = cells_body(
      columns = c(sup_1, fut_1, inf_1)
      )
    ) |>
  tab_style(
    style = list(
      cell_borders(
        sides = c("top", "bottom"), color = "red", weight = px(1), style = "solid"
      )),
    locations = list(
      cells_body(
        columns = everything(),
        rows = or_tru == "OR 1"
      )
    )
  ) |>
  tab_options(
    table.font.size = "80%"
  ) 

g_tbl
```

