---
title: "Simulation setup"

---

```{r, echo = FALSE}
source("./R/init.R")
log_info("Called simulation-pars notebook")
```



## Baseline response

The baseline probability/log-odds of treatment success is assumed to vary by silo and site of infection as detailed below.

::: {.column width="50%"}

| Silo |  Joint   | Pr(trt success) | log-odds
|------|----------|-----------|---------------|
| early |  knee   | 0.65      | 0.62              
| early |  hip    | 0.75      | 1.10
| late  |   knee  | 0.55      | 0.20
| late  |   hip   | 0.6       | 0.41
| chronic | knee  | 0.6       | 0.41
| chronic |   hip | 0.65      | 0.62

: Baseline probability of treatment success by silo and site of infection {.striped .hover}
:::

## Accrual

Accrual is assumed to follow a non-homogeneous Poisson process event times with ramp up over the first 12 months of enrolment and then enrolment of around 1.5 per day.

```{r}
#| label: fig-accrual1
#| fig-cap: 'Expected accrual'
#| fig.height: 5

# events per day
lambda = 1.52
# ramp up over 12 months 
rho = function(t) pmin(t/360, 1)

d_fig <- data.table(
  t = 0:(5 * 365),
  # expected number enrolled
  n = c(0, nhpp.mean(lambda, rho, t1 = 5 * 365, num.points = 5 * 365))
)

ggplot(d_fig, aes(x = t/365, y = n)) +
  geom_line() +
  scale_x_continuous("Year") +
  scale_y_continuous("E[Accrual]", breaks = seq(0, 2500, by = 500))
```


```{r, echo = F, eval = F}
Nsim <- 1000

lsim <- list()
for(i in 1:Nsim){
  lsim[[i]] <- c(0, nhpp.event.times(lambda, 2499, rho))
}
tt <- colMeans(do.call(rbind, lsim))

d_fig <- data.table(
  id = 1:2500,
  t0 = tt
)

ggplot(d_fig, aes(x = t0/375, y = id)) +
  geom_line() +
  scale_x_continuous("Year") +
  scale_y_continuous("Accrual", breaks = seq(0, 2500, by = 500))

```

## Domain non-membership effects

We assume a small effects for not being randomised to a domain for all domains.

## Missingness

Missingness is not implemented.

## Non-differential follow-up

To avoid artifacts associated with non-differential follow-up (e.g. early vs late deaths), participants will be included in the analyses only when they reach the primary endpoints (12 months) irrespective of whether they experienced treatment failure before that time.

## Simulation scenarios

We consider a range of simulation settings:

+ Null scenario - no effect of any randomised treatment in any cell (OR = 1) for all domains and silos
+ All effective - all randomised treatments in all cells effective (OR = 1.5) for all domains and silos
+ Single effective - single cell effective (OR = 1.5) for single domain and silo
+ Single harmful - single cell harmful (OR = 1/1.5) for single domain and silo


## Statistical quantities

We consider treatment superiority, inferiority and futility.

As a set, whatever quantities are used, need to be mutually exclusive.
For example, with the current set of quantities, if a parameter is superior, then it cannot be either inferior and futile.

### Superiority

\begin{aligned}
\phi_{sup} = Pr(OR > 1)
\end{aligned}

and conclude superiority if $\phi_{sup} > \delta_{sup}$.

### Inferiority

\begin{aligned}
\phi_{inf} = 1 - \phi_{sup}
\end{aligned}

and conclude inferiority if $\phi_{inf} > \delta_{inf}$.

### Futility

Is simply defined in relation to superiority with futility concluded if $\phi_{sup} < \delta_{fut}$.

## Decision thresholds

Decision thresholds are documented with the simulation results.








