---
title: "Simulated trial data"
date: today
date-modified: last-modified
---

```{r, echo = FALSE}
source("./R/init.R")
log_info("Called trial-data notebook")
```


Trial data is simulated under the model specification, see earlier section.
The sub-totals sum to the size of the subset (e.g. patients under late stage infection) and not the size of the total sample.
The missingness is a consequence of the partial factorial structure.

```{r}
set.seed(1)
ll <- get_trial_data(N = 2500)

d <- copy(ll$d)


```

Only patients in the late stage infection silo enter into the surgical domain for randomised comparison of surgery type.
@tbl-smry-late-Da shows the allocation to dair vs rev for this cohort and the balance across the remaining group levels in the data.

```{r, eval = T}
#| label: tbl-smry-late-Da
#| tbl-cap: 'Simulated trial data for (late silo) surgical domain - covariate balance across other groups'

# domain B
d_tmp1 <- d[l1 == 1 & er == 1, .(er, r, ed, srp, d)]
d_tmp2 <- d_tmp1[, .N, keyby = .(r, ed, srp, d)]
d_B <- dcast(d_tmp2, ed + d + srp ~ r, value.var = "N")
d_B[, domain := "Duration"]
colnames(d_B) <- c("rand", "group", "surgery", "dair", "rev", "domain")
setcolorder(d_B, "domain")
# domain C
d_tmp1 <- d[l1 == 1 & er == 1, .(er, r, ef, srp, f)]
d_tmp2 <- d_tmp1[, .N, keyby = .(r, ef, srp, f)]
d_C <- dcast(d_tmp2, ef + f + srp ~ r, value.var = "N")
d_C[, domain := "Choice"]
colnames(d_C) <- c("rand", "group", "surgery", "dair", "rev", "domain")
setcolorder(d_C, "domain")

d_tbl <- rbind(d_B, d_C, fill = TRUE)
# d_tbl[, group := factor(group, levels = c("w12","w12p1", "w06p1", "w12p2", "d07p2", "other", "norif", "rif"))]
d_tbl <- d_tbl[order(domain, rand, surgery, group)]

d_tbl[, total := rowSums(d_tbl[, .(dair, rev)], na.rm = T)]
d_tbl[domain == "Choice" & rand == 0, group := NA]
d_tbl[domain == "Duration" & rand == 0, group := NA]

cols <- c("Domain", "Revealed", "Treatment", "Surgery recvd", "DAIR", "Revision", "Total")
names(cols) <- names(d_tbl)

gt_tbl <- d_tbl |> 
  gt(groupname_col = "domain"
  )  |>
  tab_spanner(
    label = html("Domain A (late silo)"),
    columns = c(dair, rev),
    id = "da"
  ) |>
  sub_missing(
    columns = everything(),
    missing_text = "-"
  ) |>
  cols_label(.list = cols) |> 
  summary_rows(
    columns = c("dair", "rev", "total"),
    fns = list(
      subtotal = ~ sum(., na.rm = TRUE)
    )
  )  |>
  tab_footnote(
    footnote = "Revealed indicates whether units were randomised into choice/duration domain (0: No, 1: Yes)",
    locations = cells_column_labels(columns = rand)
  ) |>
  tab_footnote(
    footnote = "Surgery recvd indicates surgery type actually perfromed (0: dair, 1: one-stage, 2: two-stage). Surgery recvd may deviate from original randomisation/plan.",
    locations = cells_column_labels(columns = surgery)
  ) |>
  tab_footnote(
    footnote = "Treatment indicates which treatment assigned within domain. For choice (0: no-rif, 1: rif). For duration is specific to surgery type received (0: long, 1: short). Units that were not randomised within the choice/duration domain have undefined randomised treatment status.",
    locations = cells_column_labels(columns = group)
  ) |>
  tab_footnote(
    footnote = "Only units randomised to surgery domain reported.",
    locations = cells_column_spanners(spanners = "da")
  )

gt_tbl
```

@tbl-smry-Db shows the allocation to the duration domain conditional on surgery type actually received and the balance across the remaining group levels in the data.
Only units randomised within the duration domain are reported.

```{r, eval = T}
#| label: tbl-smry-Db
#| tbl-cap: 'Simulated trial data for AB duration domain - covariate balance across other groups'

# domain B
d_tmp1 <- d[ed == 1, .(ed, d, er, r, srp)]
d_tmp2 <- d_tmp1[, .N, keyby = .(d, er, r,srp)]
# d_tmp2[, b := factor(b, levels = c("w12", "w06p1", "w12p1", "d07p2", "w12p2"))]
d_B <- dcast(d_tmp2, er + r ~ srp + d, value.var = "N")
d_B[, domain := "Surgery"]
colnames(d_B) <- c("rand", "group", c("long_1", "short_1", "long_2", "short_2"), "domain")
setcolorder(d_B, "domain")
# domain C
d_tmp1 <- d[ed == 1, .(ed, d, ef, f, srp)]
d_tmp2 <- d_tmp1[, .N, keyby = .(d, ef, f, srp)]
# d_tmp2[, b := factor(b, levels = c("w12", "w06p1", "w12p1", "d07p2", "w12p2"))]
d_C <- dcast(d_tmp2, ef + f ~ srp + d, value.var = "N")
d_C[, domain := "Choice"]
colnames(d_C) <- c("rand", "group", c("long_1", "short_1", "long_2", "short_2"), "domain")
setcolorder(d_C, "domain")

d_tbl <- rbind(d_B, d_C, fill = T)

d_tbl[, total := rowSums(d_tbl[, .(long_1, short_1, long_2, short_2)], na.rm = T)]
d_tbl[domain == "Surgery" & rand == 0, group := NA]
d_tbl[domain == "Choice" & rand == 0, group := NA]

cols <- c("Domain", "Revealed", "Treatment", c("long", "short", "long", "short"), "Total")
names(cols) <- names(d_tbl)

gt_tbl <- d_tbl |> 
  gt(groupname_col = "domain"
  ) |> 
  summary_rows(
    columns = c("long_1", "short_1", "long_2", "short_2", "total"),
    fns = list(
      subtotal = ~ sum(., na.rm = TRUE)
    )
  )   |>
  tab_spanner(
    label = html("AB Duration <br>(one-stage)"),
    columns = c("long_1", "short_1"),
    id = "d1"
  ) |>
  tab_spanner(
    label = html("AB Duration <br>(two-stage)"),
    columns = c("long_2", "short_2"),
    id = "d2"
  ) |>
  sub_missing(
    columns = everything(),
    missing_text = "-"
  ) |>
  cols_label(.list = cols)  |>
  tab_footnote(
    footnote = "Revealed indicates whether units were randomised into surgery/duration domain (0: No, 1: Yes)",
    locations = cells_column_labels(columns = rand)
  ) |>
  tab_footnote(
    footnote = "Treatment indicates which treatment assigned within domain. For choice (0: no-rif, 1: rif). For duration is specific to surgery type received (0: long, 1: short). Units that were not randomised within the surgery/choice domain have undefined randomised treatment status.",
    locations = cells_column_labels(columns = group)
  ) |>
  tab_footnote(
    footnote = "Only units randomised to interventions in the surgery domain reported. One-stage/two-stage is by clinician selection",
    locations = cells_column_spanners(spanners = c("d1", "d2"))
  )



gt_tbl
```

@tbl-smry-Dc shows the allocation to the duration domain conditional on surgery type actually received and the balance across the remaining group levels in the data.
Only units randomised within the antibiotic choice domain are reported.


```{r, eval = T}
#| label: tbl-smry-Dc
#| tbl-cap: 'Simulated trial data for AB choice domain - covariate balance across other groups'

# domain 
d_tmp1 <- d[ef == 1, .(f, er, r, srp)]
d_tmp2 <- d_tmp1[, .N, keyby = .(f, er, r, srp)]
# d_tmp2[, a := factor(a, levels = c("dair", "rev", "one", "two"))]
# d_tmp2[, qa := factor(qa, levels = c("dair", "one", "two"))]
# d_tmp2[, c := factor(c, levels = c("other", "norif", "rif"))]
d_A <- dcast(d_tmp2, er + srp + r ~ f, value.var = "N")
d_A[, domain := "Surgery"]
colnames(d_A) <- c("rand", "surgery","group",  c("norif", "rif"), "domain")
setcolorder(d_A, "domain")
# domain
d_tmp1 <- d[ef == 1, .(f, ed, d, srp)]
d_tmp2 <- d_tmp1[, .N, keyby = .(f, ed, d, srp)]
# d_tmp2[, b := factor(b, levels = c("w12", "w06p1", "w12p1", "d07p2", "w12p2"))]
# d_tmp2[, c := factor(c, levels = c("other", "norif", "rif"))]
d_B <- dcast(d_tmp2, ed + srp + d ~ f, value.var = "N")
d_B[, domain := "Duration"]
colnames(d_B) <- c("rand", "surgery", "group", c("norif", "rif"), "domain")
setcolorder(d_B, "domain")

d_tbl <- rbind(d_A, d_B, fill = T)

d_tbl[, total := rowSums(d_tbl[, .(norif, rif)], na.rm = T)]

d_tbl[domain == "Surgery" & rand == 0, group := NA]
d_tbl[domain == "Duration" & rand == 0, group := NA]



cols <- c("Domain", "Revealed", "Surgery recvd", "Treatment", c("norif", "rif"), "Total")
names(cols) <- names(d_tbl)

gt_tbl <- d_tbl |> 
  gt(groupname_col = "domain"
  ) |> 
  summary_rows(
    columns = c("norif", "rif", "total"),
    fns = list(
      subtotal = ~ sum(., na.rm = TRUE)
    )
  ) |>
  tab_spanner(
    label = html("AB Choice"),
    columns = c("norif", "rif"),
    id = "c1"
  ) |>
  sub_missing(
    columns = everything(),
    missing_text = "-"
  ) |>
  cols_label(.list = cols)|>
  tab_footnote(
    footnote = "Revealed indicates whether units were randomised into surgery/choice domain (0: No, 1: Yes)",
    locations = cells_column_labels(columns = rand)
  ) |>
  tab_footnote(
    footnote = "Treatment indicates which treatment assigned within domain. For surgery (0: dair, 1: rev). For duration is specific to surgery type received (0: long, 1: short). Units that were not randomised within the surgery/duration domain have undefined randomised treatment status.",
    locations = cells_column_labels(columns = group)
  ) |>
  tab_footnote(
    footnote = "Only units randomised to interventions in the antibiotic choice domain reported.",
    locations = cells_column_spanners(spanners = c("c1"))
  )

gt_tbl
```



